\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2019-01-14 redaction/old/note_secondstage_v1512.tex
rangement dossier redaction + coquille intro
creation dossier "old" pour vieille version + correction reference manquante en intro (Lashkaripour 2018, on avait un beau "?" en gras au beau milieu dela page 4 ;-))

\paragraph{Insurance}: Based on information collected about insurance grids \textbf{(source???)}\footnote{\textbf{See Appendix ??? for details.}}, we infer that insurance costs depend on the value of the shipment, conditional on the type of product transported (Insuring one ton of diamonds does not bear the same cost as a ton of wood). The database provides us with the quantity (in terms of kilograms) of the shipment ($q_{ik}$), as well as the price, so that we have the value of the shipment. However, we have no dedicated variable to measure the insurance cost in the database. Accordingly, we capture this dimension by including product fixed effects. Insurance grids retain a degree of product categorization much broader than the 3-digit level we retain as benchmark sectorial degree (our $k$ index for the product dimension). In most pieces of information we have collected, insurance grids adopt a ranking in \textbf{20 (??)} types of products. Accordingly, we include product fixed effects of dimension \textbf{20}, that we refer to with the subscript $k^\prime$, i.e. $FE_{k^\prime}$. If we denote $p_{ik}^{fob}$ and $q_{ik}$ the (fas) price and the quantity of the shipment, total insurance costs (in USD) are accordingly measured by $FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}$. As insurance cost depend on the value of the shipment, we expect it to be proportional to the price, ie rather play on the iceberg part of the transport costs.

\paragraph{Costs} (for handling costs). The amount paid to load the bulk on transport mode is likely to depend on the quantity of product charged, as well as the bulk implied in the shipping, that is referred to as the volumetric weight (A ton of cars does not represent the same volume on board as a ton of apparel, and accordingly does not imply the same handling costs); further, handling costs also depend on the country of origin (quality of infrastructure, labor and various other costs). If the database provides us with the quantity that is imported (of product $k$ from country $i$), we do not have information about the volume it takes, which affects the handling costs. We capture this dimension using information provided by the \textit{Maritime Transport Costs} database provided by the OECD. As detailed in Appendix \ref{app:data}, we use this database to build a measure of the product-specific cost to export at the 3-digit level, that we use as proxy for the handling costs specific to the volumetric weight of products (denoted by $V_{k}$). Last, handling costs are likely to be country-dependent (in the quality of infrastructure in particular). We capture this dimension with the ``Cost to export'' variable provided by the World Bank (see Appendix \ref{app:data} for details), denoted $EC_{i}$ in the following. Accordingly, we measure total handling costs (in USD) by $EC_i\times V_{k}\times q_{ik}$. As such costs depend on the quantity that is shipped, but not on the price, we expect this dimension to play on the additive part of transport costs.

 \paragraph{Freight}. Quite intuitively, freight costs should depend on the distance between the origin country and the U.S. (denoted $dist_i$). Further, they likely depend on the quantity ($q_{ik}$) and the bulk of the shipment. Again, we capture this dimension through product fixed effects, based on the same product categorization (based on $k^\prime$) as for handling and insurance costs. Accordingly, we capture freight costs through $V_{k}\times dist_i\times q_{ik}$. Because this cost is most likely dependent on the quantity of goods exported rather than their value, we expect this term to have more impact on the additive part of total transport costs.

@@

with $\varepsilon_{ik}$ the error term, centered on 0. This calls for two comments. First, we add the country fixed effects separately in the regression, in order to correctly identify the associated interaction term. Second, we specify the error term multiplicatively. If we estimate the equation by setting the error term in addition to the other variables, it means that the error term is expressed as a percentage of the fob price (ie, it is independent of the other dimensions, the residual term can be interpreted in percentage points). In this case, it is very likely that the variance of the error term depends on the size of the observed costs. E.g., because costs and their measures are bounded by zero, the error has to be small when costs are small and it might be higher when costs are high. Adopting this functional form would then be subject to a potential bias in point estimates due to the heteroscedasticity of error terms. This drives us to specify the error term in a multiplicative manner (such that the error term is in percentage of each explicative variable and centered on 0), and to take the log of the equation.

@@

%\textbf{Careful:} Again, challenging the role of distance (as we ambition) may receive limited support, in that the counter-argument we should expect: ``ok it has no role on the iceberg part of transport costs, and all on additive component of transport costs, but you don't show that it does not play on the other dimensions of trade costs you don't cover (on top of the cif-fob difference) multiplicatively. So, you don't demonstrate what you argue, that distance is not a good proxy for trade costs.'' ... \bigskip

%What we are doing here is descriptive statistics exercise; therefore, it is normal that we do not report standard deviations !

\subsection{Results and Interpretation}
%\section{One step further into interpretation: transportation costs as an indicator of market structure.}

\section{Conclusion}

\newpage
\bibliographystyle{essaien}
\bibliography{biblio}

\newpage

@@

\section*{Data Appendix \label{app:data}}

@@

Doing Business measures the time and cost (excluding tariffs) associated with exporting and importing a standardized cargo of goods by sea transport. The most recent round of data collection for the project was completed in June 2014. Among the four variables provided, we retain the variable ``Cost to export'' (US\$ per container), as it represents the ``cost associated with all procedures required to export goods[...,] including the costs for documents, administrative fees for customs clearance and technical control, customs broker fees, terminal handling charges and inland transport.'' We use the variable observed \textbf{in year ??? attention la variable est-elle destination spefici? A priori non, a verifier.}
\newpage

\section*{Tables and Figures}

%%%Figures
% Goodness-of-fit

\begin{figure}[htbp]
\caption{Multiplicative Costs (-1), Median Value 3 digits}\label{fig:good_fit}
\begin{center}
\begin{tabular}{cc}
{\small (a) Air } & {\small (b) Vessel}\\
\includegraphics[width=3.5in, height=3in]{graph1a.pdf}
& \includegraphics[width=3.5in,height=3in]{graph1b.pdf} \\
\end{tabular}
\end{center}
\end{figure}

%Economic interpretation
%usepackage{float} + H is great to keep images on the right page
\begin{figure}[H]
\caption{Multiplicative vs. Additive Trade Costs, Median Value 3 digits}\label{fig:mult_vs_add}
\begin{center}
\begin{tabular}{cc}
{\small (a) Air } & {\small (b) Vessel}\\
\includegraphics[width=3.5in, height=3in]{graph2a.pdf}
& \includegraphics[width=3.5in,height=3in]{graph2b.pdf} \\
\end{tabular}
\end{center}
\end{figure}

%%%%Tables
%Goodness-of-fit
\begin{table}[htbp]
  \centering
  \caption{Air: Measures of Goodness-of-fit}
\begin{center}
    \begin{tabular}{lccccccc}
  %  \toprule
   \hline\hline
          & \multicolumn{6}{c}{Year}                      &  \\

      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1980} & \multicolumn{1}{c}{1990} & \multicolumn{1}{c}{2000} & \multicolumn{1}{c}{2010} & \multicolumn{1}{c}{2013} & \multicolumn{1}{c}{\textbf{Mean stat}}\\
\hline
\textbf{R$^{2}$} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
Term I only & 0.30  & 0.27  & 0.25  & 0.32  & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.31} & \multicolumn{1}{c}{\textbf{0.31}} \\
Terms A \& I & 0.59  & 0.65  & 0.63  & 0.64  & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{\textbf{0.60}} \\
\textbf{SER} &       &       &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
Term I only & 0.79  & 0.86  & 0.81  & 0.84  & \multicolumn{1}{c}{0.86} & \multicolumn{1}{c}{0.92} & \multicolumn{1}{c}{\textbf{0.85}} \\
Terms A \& I & 0.67  & 0.71  & 0.67  & 0.70  & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.85} & \multicolumn{1}{c}{\textbf{0.73}} \\
\textbf{Log-likelihood} &       &       &       &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
Term I only & -17530.49 & -20253.49 & -29977.79 & -43341.27 & \multicolumn{1}{c}{-50746.83} & \multicolumn{1}{c}{-52690.29} & \multicolumn{1}{c}{\textbf{-34888.64}} \\
Terms A \& I & -15125.65 & -17263.20 & -25393.46 & -36788.44 & \multicolumn{1}{c}{-47277.53} & \multicolumn{1}{c}{-49419.70} & \multicolumn{1}{c}{\textbf{-30508.29}} \\
\textbf{AIC criteria} &       &       &       &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
Term I only & 35674.98 & 41170.98 & 60715.58 & 87492.55 & \multicolumn{1}{c}{102297.66} & \multicolumn{1}{c}{106130.58} & \multicolumn{1}{c}{\textbf{70498.08}} \\
Terms A \& I & 31387.29 & 35738.39 & 52098.91 & 74954.88 & \multicolumn{1}{c}{95887.05} & \multicolumn{1}{c}{100155.41} & \multicolumn{1}{c}{\textbf{62284.99}} \\
\textbf{Test LL} &       &       &       &       & \multicolumn{1}{c}{} &       &  \\
2$\times$(ll(UR) -ll(R)) & 4809.68 & 5980.59 & 9168.67 & 13105.67 & \multicolumn{1}{c}{6938.61} & \multicolumn{1}{c}{6541.17} & \multicolumn{1}{c}{\textbf{8760.69}} \\
\# restrictions  & 355   & 369   & 393   & 426   & \multicolumn{1}{c}{426.00} & \multicolumn{1}{c}{422} & \multicolumn{1}{c}{\textbf{401.93}} \\
p-value & 0.000 & 0.000 & 0.000 & 0.000 & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{\textbf{}} \\

\hline\hline
   % \bottomrule

    \end{tabular}%
    \end{center}
  \label{tab:good_fit_air}%
  \tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of regression; AIC = Akaike Information Criterion. Statistics are obtained weighting each observation by its value in transport (mode-dependent). Term A expressed in fraction of fob price. R$^{2}$ between the log of predicted ratio and the log of the observed ratio. The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
\end{table}%

\begin{table}[htbp]
  \centering
  \caption{Vessel, Measures of Goodness-of-fit}
\begin{center}
    \begin{tabular}{lccccccc}
  %  \toprule
   \hline\hline
          & \multicolumn{6}{c}{Year}                      &  \\
          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \textbf{Mean stat} \\
\hline
\textbf{R$^{2}$}\unboldmath{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       &  \\
Term I only & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{0.40} & \multicolumn{1}{c}{0.35} & \multicolumn{1}{c}{0.34} & \multicolumn{1}{c}{\textbf{0.39}} \\
Terms A \& I & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.57} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{\textbf{0.56}} \\
\textbf{SER} &       &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
Term I only & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.74} & \multicolumn{1}{c}{0.76} & \multicolumn{1}{c}{\textbf{0.66}} \\
Terms A \& I & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.55} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{\textbf{0.57}} \\
\textbf{Log-likelihood} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
Term I only & \multicolumn{1}{c}{-16287.40} & \multicolumn{1}{c}{-16129.13} & \multicolumn{1}{c}{-25169.31} & \multicolumn{1}{c}{-35263.95} & \multicolumn{1}{c}{-41998.95} & \multicolumn{1}{c}{-43692.93} & \multicolumn{1}{c}{\textbf{-28534.30}} \\
Terms A\& I & \multicolumn{1}{c}{-12985.76} & \multicolumn{1}{c}{-13353.65} & \multicolumn{1}{c}{-21171.37} & \multicolumn{1}{c}{-29490.96} & \multicolumn{1}{c}{-37418.66} & \multicolumn{1}{c}{-39751.86} & \multicolumn{1}{c}{\textbf{-24151.31}} \\
\textbf{AIC criteria} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{\textbf{}} \\
Term I only & \multicolumn{1}{c}{33328.81} & \multicolumn{1}{c}{33010.27} & \multicolumn{1}{c}{51142.62} & \multicolumn{1}{c}{71365.89} & \multicolumn{1}{c}{84789.89} & \multicolumn{1}{c}{88191.87} & \multicolumn{1}{c}{\textbf{57848.60}} \\
Terms A \& I & \multicolumn{1}{c}{27331.52} & \multicolumn{1}{c}{28067.31} & \multicolumn{1}{c}{43664.74} & \multicolumn{1}{c}{60475.91} & \multicolumn{1}{c}{76161.33} & \multicolumn{1}{c}{80873.72} & \multicolumn{1}{c}{\textbf{49683.32}} \\
\textbf{Test LL} &       &       &       &       & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{\textbf{}} \\
2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{6603.28} & \multicolumn{1}{c}{5550.96} & \multicolumn{1}{c}{7995.88} & \multicolumn{1}{c}{11545.98} & \multicolumn{1}{c}{9160.56} & \multicolumn{1}{c}{7882.15} & \multicolumn{1}{c}{\textbf{8765.97}} \\
\# restrictions  & \multicolumn{1}{c}{393} & \multicolumn{1}{c}{395} & \multicolumn{1}{c}{411} & \multicolumn{1}{c}{436} & \multicolumn{1}{c}{424} & \multicolumn{1}{c}{427} & \multicolumn{1}{c}{\textbf{417.05}} \\
p-value & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\

\hline\hline
   % \bottomrule

    \end{tabular}%
    \end{center}
  \label{tab:good_fit_ves}%
  \tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of Regression; AIC = Akaike Information Criterion. Statistics are obtained weighting each observation by its value in transport (mode-dependent). Term A expressed in fraction of fob price. R$^{2}$ between the log of predicted ratio and the log of the observed ratio. The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
\end{table}%

%Economic Interpretation

\begin{table}[htbp]
  \centering
  \caption{Air: Dynamics of Trade Costs over Time}
\begin{center}
    \begin{tabular}{lccccccc}

  %  \toprule
   \hline\hline
          & \multicolumn{6}{c}{Year}                      &  \\

          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \multicolumn{1}{c}{\textbf{Mean stat}} \\
 \hline
\multicolumn{1}{l}{\textbf{With only Iceberg Trade Costs}} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
Mean  & 1.069 & 1.054 & 1.050 & 1.036 & 1.042 & 1.034 & \multicolumn{1}{c}{\textbf{1.051}} \\
Median & 1.054 & 1.038 & 1.044 & 1.025 & 1.034 & 1.029 & \multicolumn{1}{c}{\textbf{1.042}} \\
Standard Error & 0.052 & 0.049 & 0.039 & 0.033 & 0.037 & 0.024 & \multicolumn{1}{c}{\textbf{0.042}}\\
\hline
\multicolumn{1}{l}{\textbf{With Additive \& Iceberg Trade Costs }} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}}\\
\multicolumn{1}{l}{\textit{Additive term }} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
\multicolumn{1}{l}{Mean } & 0.026 & 0.020 & 0.018 & 0.013 & 0.011 & 0.010 & \multicolumn{1}{c}{\textbf{0.019}} \\
\multicolumn{1}{l}{Median} & 0.011 & 0.005 & 0.008 & 0.005 & 0.004 & 0.005 & \multicolumn{1}{c}{\textbf{0.007}} \\
\multicolumn{1}{l}{Standard Error} & 0.040 & 0.041 & 0.033 & 0.028 & 0.024 & 0.020 & \multicolumn{1}{c}{\textbf{0.076}} \\
\multicolumn{1}{l}{\textit{Iceberg term}} &       &       &       &       &       &       & \multicolumn{1}{c}{\textbf{}} \\
\multicolumn{1}{l}{Mean } & 1.036 & 1.023 & 1.024 & 1.017 & 1.026 & 1.017 & \multicolumn{1}{c}{\textbf{1.025}} \\
\multicolumn{1}{l}{Median} & 1.027 & 1.016 & 1.016 & 1.012 & 1.022 & 1.017 & \multicolumn{1}{c}{\textbf{1.018}} \\
Standard Error & 0.032 & 0.025 & 0.021 & 0.016 & 0.023 & 0.012 & \multicolumn{1}{c}{\textbf{0.023}} \\
\hline
\# observations & 14955 & 16118 & 24958 & 35027 & 40279 & 39351 & \textbf{} \\

\hline\hline
    \end{tabular}%
  \end{center}
  \label{tab:result_air}%
\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
\end{table}%

\begin{table}[htbp]
  \centering
  \caption{Vessel: Dynamics of Trade Costs over Time}
\begin{center}
    \begin{tabular}{lccccccc}

  %  \toprule
   \hline\hline
          & \multicolumn{6}{c}{Year}                      &  \\

          & 1974  & 1980  & 1990  & 2000  & 2010  & 2013  & \multicolumn{1}{c}{\textbf{Mean stat}} \\
 \hline
   \textbf{With only Iceberg Trade Costs} &       &       &       &       &       &       & \textbf{} \\
Mean  & 1.098 & 1.065 & 1.057 & 1.051 & 1.040 & 1.036 & \textbf{1.058} \\
Median & 1.096 & 1.055 & 1.046 & 1.049 & 1.036 & 1.033 & \textbf{1.051} \\
Standard Error & 0.053 & 0.040 & 0.032 & 0.028 & 0.020 & 0.018 & \textbf{0.032} \\
\textbf{With Additive \& Iceberg Trade Costs} &       &       &       &       &       &       & \textbf{} \\
\hline
\textit{Additive term } &       &       &       &       &       &       & \textbf{} \\
Mean  & 0.051 & 0.034 & 0.027 & 0.028 & 0.025 & 0.015 & \textbf{0.029} \\
Median & 0.029 & 0.023 & 0.017 & 0.022 & 0.019 & 0.008 & \textbf{0.019} \\
Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 & \textbf{0.041} \\
\textit{Iceberg term} &       &       &       &       &       &       & \textbf{} \\
Mean  & 1.054 & 1.031 & 1.033 & 1.025 & 1.019 & 1.022 & \textbf{1.032} \\
Median & 1.049 & 1.024 & 1.028 & 1.021 & 1.018 & 1.018 & \textbf{1.028} \\
Standard Error & 0.041 & 0.023 & 0.022 & 0.021 & 0.018 & 0.018 & \textbf{0.028} \\
\hline
\# obs & 19007 & 17356 & 28383 & 36090 & 37748 & 38473 & \textbf{} \\

\hline\hline
    \end{tabular}%
  \end{center}
  \label{tab:result_ves}%
\tiny{Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fob price. }
\end{table}%

\newpage

\section{Results with a Higher Level of Products Disaggregation }
%\vspace{-0.5cm}
\begin{table}[htbp]
  \centering
  \caption{Air: Measures of Goodness-of-fit, 4-digits}
\begin{center}
\label{tab:good_fit_air_rob}%
%\vspace{0.5cm}
\scalebox{0.97}{
\begin{tabular}{lccccccc}
\hline
\hline
      & \multicolumn{6}{c}{Year}                      &  \\
      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013} & \multicolumn{1}{c}{\textbf{Mean stat}} \\
\hline
\textbf{R2} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       &       &       & \multicolumn{1}{c}{\textbf{}}\\
Term I only & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} & \textbf{0.47} \\
Terms A \& I & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} & \textbf{0.63} \\
\textbf{SER} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \textbf{} \\
Term I only & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93} & \textbf{0.89} \\
Terms A \& I & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86} & \textbf{0.79} \\
\textbf{Log-likelihood} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \textbf{} \\
Term I only & \multicolumn{1}{c}{-17505.55} & \multicolumn{1}{c}{-21813.46} & \multicolumn{1}{c}{-30960.56} & \multicolumn{1}{c}{-44067.62} & \multicolumn{1}{c}{-49375.57} & \multicolumn{1}{c}{-53197.87} & \textbf{-34744.40} \\
Terms A\& I & \multicolumn{1}{c}{-14895.81} & \multicolumn{1}{c}{-18589.91} & \multicolumn{1}{c}{-26553.53} & \multicolumn{1}{c}{-37297.93} & \multicolumn{1}{c}{-45747.57} & \multicolumn{1}{c}{-49899.14} & \textbf{-30243.95} \\
\textbf{AIC criteria} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \textbf{} \\
Term I only & \multicolumn{1}{c}{36243.10} & \multicolumn{1}{c}{44966.91} & \multicolumn{1}{c}{63417.12} & \multicolumn{1}{c}{89747.24} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.73} & \textbf{70940.07} \\
Terms A \& I & \multicolumn{1}{c}{31873.63} & \multicolumn{1}{c}{39495.82} & \multicolumn{1}{c}{55777.05} & \multicolumn{1}{c}{77439.85} & \multicolumn{1}{c}{94059.14} & \multicolumn{1}{c}{102224.28} & \textbf{62955.73} \\
\textbf{Test LL} &       &       &       &       &       &       & \textbf{} \\
2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{5219.47} & \multicolumn{1}{c}{6447.09} & \multicolumn{1}{c}{8814.06} & \multicolumn{1}{c}{13539.39} & \multicolumn{1}{c}{7255.99} & \multicolumn{1}{c}{6597.45} & \textbf{9000.89} \\
\# restrictions  & \multicolumn{1}{c}{640} & \multicolumn{1}{c}{698} & \multicolumn{1}{c}{778} & \multicolumn{1}{c}{833} & \multicolumn{1}{c}{824} & \multicolumn{1}{c}{818} & \textbf{755.73} \\
p-value & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \textbf{} \\
\hline\hline
\end{tabular}}%
\end{center}
 \vspace{-10pt} \tiny{Note: I = Iceberg; A = Additive; SER = Standard Error of regression; AIC = Akaike Information Criterion. Statistics are obtained weighting each observation by its value in transport (mode-dependent). Term A expressed in fraction of fob price. R$^{2}$ between the log of predicted ratio and the log of the observed ratio. The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}
\end{table}%

\begin{table}[H]
  \centering
  \caption{Vessel: Measures of Goodness-of-fit, 4-digits}
\begin{center}
\label{tab:good_fit_ves_rob}%

>>>>

\begin{itemize}
\item Insurance: Based on information collected about insurance grids \textbf{(source???)}\footnote{\textbf{See Appendix ??? for details.}}, we infer that insurance costs depend on the value of the shipment, conditional on the type of product transported (Insuring one ton of diamonds does not bear the same cost as a ton of wood). The database provides us with the quantity (in terms of kilograms) of the shipment ($q_{ik}$), as well as the price, so that we have the value of the shipment. However, we have no dedicated variable to measure the insurance cost in the database. Accordingly, we capture this dimension by including product fixed effects. Insurance grids retain a degree of product categorization much broader than the 3-digit level we retain as benchmark sectorial degree (our $k$ index for the product dimension). In most pieces of information we have collected, insurance grids adopt a ranking in \textbf{20 (??)} types of products. Accordingly, we include product fixed effects of dimension \textbf{20}, that we refer to with the subscript $k^\prime$, i.e. $FE_{k^\prime}$. If we denote $p_{ik}^{fob}$ and $q_{ik}$ the (fas) price and the quantity of the shipment, total insurance costs (in USD) are accordingly measured by $FE_{k^\prime}\times p_{ik}^{fob}\times q_{ik}$. As insurance cost depend on the value of the shipment, we expect it to be proportional to the price, ie rather play on the iceberg part of the transport costs.

\item Costs (for handling costs). The amount paid to load the bulk on transport mode is likely to depend on the quantity of product charged, as well as the bulk implied in the shipping, that is referred to as the volumetric weight (A ton of cars does not represent the same volume on board as a ton of apparel, and accordingly does not imply the same handling costs); further, handling costs also depend on the country of origin (quality of infrastructure, labor and various other costs). If the database provides us with the quantity that is imported (of product $k$ from country $i$), we do not have information about the volume it takes, which affects the handling costs. We capture this dimension using information provided by the \textit{Maritime Transport Costs} database provided by the OECD. As detailed in Appendix \ref{app:data}, we use this database to build a measure of the product-specific cost to export at the 3-digit level, that we use as proxy for the handling costs specific to the volumetric weight of products (denoted by $V_{k}$). Last, handling costs are likely to be country-dependent (in the quality of infrastructure in particular). We capture this dimension with the ``Cost to export'' variable provided by the World Bank (see Appendix \ref{app:data} for details), denoted $EC_{i}$ in the following. Accordingly, we measure total handling costs (in USD) by $EC_i\times V_{k}\times q_{ik}$. As such costs depend on the quantity that is shipped, but not on the price, we expect this dimension to play on the additive part of transport costs.

\item Freight. Quite intuitively, freight costs should depend on the distance between the origin country and the U.S. (denoted $dist_i$). Further, they likely depend on the quantity ($q_{ik}$) and the bulk of the shipment. Again, we capture this dimension through product fixed effects, based on the same product categorization (based on $k^\prime$) as for handling and insurance costs. Accordingly, we capture freight costs through $V_{k}\times dist_i\times q_{ik}$. Because this cost is most likely dependent on the quantity of goods exported rather than their value, we expect this term to have more impact on the additive part of total transport costs.

\end{itemize}

@@

with $\varepsilon_{ik}$ the error term, centered on 0. This calls for two comments. First, we add the country fixed effects seperately in the regression, in order to correctly identify the associated interaction term. Second, we specify the error term multiplicatively. If we estimate the equation by setting the error term in addition to the other variables, it means that the error term is expressed as a percentage of the fob price (ie, it is independent of the other dimensions, the residual term can be interpreted in percentage points). In this case, it is very likely that the variance of the error term depends on the size of the observed costs. E.g., because costs and their measures are bounded by zero, the error has to be small when costs are small and it might be higher when costs are high. Adopting this functional form would then be subject to a potential bias in point estimates due to the heteroscedasticity of error terms. This drives us to specify the error term in a multiplicative manner (such that the error term is in percentage of each explicative variable and centered on 0), and to take the log of the equation.

@@

\textbf{Careful:} Again, challenging the role of distance (as we ambition) may receive limited support, in that the counter-argument we should expect: ``ok it has no role on the iceberg part of transport costs, and all on additive component of transport costs, but you don't show that it does not play on the other dimensions of trade costs you don't cover (on top of the cif-fob difference) multiplicatively. So, you don't demonstrate what you argue, that distance is not a good proxy for trade costs.'' ... \bigskip

\textbf{Questions raised} \medskip
\begin{itemize}
\item The literature commonly uses gravity variables (such as distance, common language, colonial linkages, common border, etc.) as a proxy for trade costs. In the second-step estimation, we only consider distance. This choice relies on the fact that we explain the transport gap measured by the gap between the fas price and the caf price. Our measure of transport cost is only part of the larger size of trade costs, and there is no a priori reason why the other gravity variables except distance might intervene in the difference between the price at the export departure point and the import arrival point.

\item Make a difference between the estimated additive trade cost and the multiplicative trade costs. Ie, should we suppose that the explicative variables enter additively when considering the additive trade costs, and multiplicatively regarding the iceberg component? No a priori reason. In both cases, the dependent variable is expressed as a percentage (of the fob price). We decompose this percentage in various dimension which, all added, compose the overall trade cost component. This justifies an additive specification, for both the additive and the iceberg trade costs components.

\end{itemize}

@@

\section{Data Appendix \label{app:data}}

@@

Doing Business measures the time and cost (excluding tariffs) associated with exporting and importing a standardized cargo of goods by sea transport. The most recent round of data collection for the project was completed in June 2014. Among the four variables provided, we retain the variable ``Cost to export'' (US\$ per container), as it represents the ``cost associated with all procedures required to export goods[...,] including the costs for documents, administrative fees for customs clearance and technical control, customs broker fees, terminal handling charges and inland transport.'' We use the variable observed \textbf{in year ??? attention la variable est-elle destination spefici? A priori non, a verifier.}
